Multi-target detection with application to cryo-electron microscopy

### 2019-03-05 inverse_problems/inverse_problems.tex
Small changes to Section 4.1 + related proofs

\begin{proposition} \label{prop:gamma}
	Assume the separation condition~\eqref{eq:spacing} holds and $K=1$ (all signal occurrences in $y$ are identical, up to noise). If the mean of $x$ is nonzero, then  
	\begin{equation}
	\gamma  = \frac{L (a^1_y)^2}{a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\sigma^2}.
	\end{equation}
\end{proposition}

Using third-order autocorrelation information of $y$, both the ratio $\gamma$ and the noise $\sigma$ are determined simultaneously. For the following results, when we say that a result holds for a ``generic'' signal $x$, we mean that the set of signals which cannot be determined by these measurements
has Lebesgue measure zero. 
In particular, this means that we can recover
almost all signals with the given measurements. The proof is provided in Appendix~\ref{sec:proof_prop_gamma_sigma}.

@@

Recall that $a_y^1 = \beta(\one^Tx)$ and $a_y^2[0] = \beta\|x\|^2+\sigma^2$, where $\one\in\RL$ is the vector of all-ones.  Consider the product:
\begin{equation}\label{eq:E1}
\begin{split}
E_1 := a_y^1a_y^2[0] =  (\beta(\one^Tx))(\beta\|x\|^2+\sigma^2)  = \sigma^2a_y^1 + L\beta^2F_1,
\end{split}
\end{equation}
where $F_1 := a_x^3[0,0] + \sum_{j=1}^{L-1}(a_x^3[j,j] + a_x^3[0,j])$. 
% Nicolas: the expression for F_1 is correct, see identifiability_of_gamma_and_sigma_Sep4_2018.m and notes Sep 4, 2018
The terms of $F_1$ can also be estimated from $a_y^3$, while taking the scaling and bias terms into account. This yields another observable:
\begin{align} 
E_2 & := a_y^3[0,0] + \sum_{j=1}^{L-1}(a_y^3[j,j] + a_y^3[0,j]) \nonumber\\
& = L\beta F_1 + (2L+1)\sigma^2a_y^1. \label{eq:E2}
\end{align}
% Nicolas: checked this one too.
Therefore, from~\eqref{eq:E1} and~\eqref{eq:E2} we get:
\begin{equation} \label{eq:E12}
E_2\beta -(2L+1)\sigma^2\beta a_y^1 = E_1-\sigma^2a_y^1.
\end{equation}
Let $E_3:=a_y^2[0] + 2\sum_{j = 1}^{L-1}a_y^2[j]$; recall from Proposition~\ref{prop:gamma}:
\begin{equation} \label{eq:sigma2}
\sigma^2 = E_3 - (a^1_y)^2/\beta. 
\end{equation} 
Plugging into~\eqref{eq:E12} and rearranging, we get a first quadratic equation in $\beta$,
\begin{equation} \label{eq:quad1}
\mathcal{A}\beta^2 + \mathcal{B}\beta + \mathcal{C} = 0,
\end{equation}
where 
\begin{align*}
\mathcal{A} &= E_2 - (2L+1)a_y^1E_3, \\ 
\mathcal{B} &= -E_1 + (2L+1)(a_y^1)^3 + a_y^1E_3  , \\
\mathcal{C} &= -(a_y^1)^3.
\end{align*}
Importantly, these coefficients are observable quantities. As we assume throughout this proof that $x$ has nonzero mean, $a_y^1 \neq 0$ and we conclude that this equation is non-trivial.

Next, we derive the second quadratic equation for $\beta$. We notice that 
\begin{equation} \label{eq:E3}
E_4 := \frac{1}{L}(a_y^1)^3 = \frac{1}{L}\beta^3 (\one ^Tx)^3   = \beta^3 F_2,
\end{equation}
where $F_2 = \frac{1}{L}(\one ^Tx)^3$, and we can work out that:
\begin{equation*}
F_2 = a_x^3[0,0] + 3\sum_{j=1}^{L-1} \left(a_x^3[j,j] + a_x^3[0,j]\right) + 6\sum_{1\leq i < j\leq L-1}a_x^3[i,j].
\end{equation*}
Once again, $F_2$ can be estimated from $a_y^3$, taking bias and scaling into account:
\begin{align}
E_5 & := a_y^3[0,0] + 3\sum_{j=1}^{L-1} \left(a_y^3[j,j] + a_y^3[0,j]\right) + 6\sum_{1\leq i < j\leq L-1}a_y^3[i,j]  = L \beta F_2 + (6L-3)\sigma^2a_y^1.
\end{align}
Consider the following ratio:
\begin{equation*} 
\frac{E_5}{E_4} = \frac{L}{\beta^2} + \frac{(6L-3)\sigma^2a_y^1}{E_4}.
\end{equation*}
From the latter, we deduce:
\begin{equation*}
\sigma^2 = \frac{E_5}{a_y^1(6L-3)}  - \frac{LE_4}{\beta^2a_y^1(6L-3)}.
\end{equation*}
Using~\eqref{eq:sigma2} and rearranging, we get the second quadratic:
\begin{equation} \label{eq:quad2}
\mathcal{D}\beta^2 + \mathcal{E}\beta + \mathcal{F} = 0,
\end{equation}
where
\begin{align*}
\mathcal{D} &= E_3 - \frac{E_5}{a_y^1(6L-3)}, \\ 
\mathcal{E} &= -(a_y^1)^2, \\
\mathcal{F} &= \frac{LE_4}{a_y^1(6L-3)}.
\end{align*}
It is also non-trivial since $E_4 \neq 0$.

>>>>

\begin{proposition} \label{prop:gamma}
	Assume the separation condition~\eqref{eq:spacing} holds and $K=1$ (all signal occurrences in $y$ are identical, up to noise). If the mean of $x$ is nonzero, then  
	\begin{equation}
	\gamma  = \frac{L (a^1_y)^2}{a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]-\sigma^2},
	\end{equation}
	meaning $\gamma$ can be determined from $\sigma$ (and vice versa) using the observables $a_y^1, a_y^2$.
\end{proposition}

Using third-order autocorrelation information of $y$, both the ratio $\gamma$ and the noise $\sigma$ can be determined simultaneously. For the following results, when we say that a result holds for a ``generic'' signal $x$, we mean that the set of signals which cannot be determined by these measurements
has Lebesgue measure zero. 
In particular, this means that we can recover
almost all signals with the given measurements. The proof is provided in Section~\ref{sec:proof_prop_gamma_sigma}.

@@

Recall that $a_y^1 = \beta(\one^Tx)$ and $a_y^2[0] = \beta\|x\|^2+\sigma^2$, where $\one\in\RL$ is the vector of all-ones and $\|x\|=\sqrt{x[0]^2 + \cdots + x[L-1]^2}$ is the 2-norm. Consider the product $E_1$:
\begin{equation}\label{eq:E1}
\begin{split}
E_1 = a_y^1a_y^2[0] =  (\beta(\one^Tx))(\beta\|x\|^2+\sigma^2)  = \sigma^2a_y^1 + L\beta^2F_1,
\end{split}
\end{equation}
where $F_1 = a_x^3[0,0] + \sum_{j=1}^{L-1}(a_x^3[j,j] + a_x^3[0,j])$. 
% Nicolas: the expression for F_1 is correct, see identifiability_of_gamma_and_sigma_Sep4_2018.m and notes Sep 4, 2018
The terms of $F_1$ can also be estimated from $a_y^3$, while taking the scaling and bias terms into account. This yields another observable, $E_2$:
\begin{align} 
E_2 & = a_y^3[0,0] + \sum_{j=1}^{L-1}(a_y^3[j,j] + a_y^3[0,j]) = L\beta F_1 + (2L+1)\sigma^2a_y^1. \label{eq:E2}
\end{align}
% Nicolas: checked this one too.
Therefore, from~\eqref{eq:E1} and~\eqref{eq:E2} we get:
\begin{equation} \label{eq:E12}
E_2\beta -(2L+1)\sigma^2\beta a_y^1 = E_1-\sigma^2a_y^1.
\end{equation}
Let $E_3 =a_y^2[0] + 2\sum_{\ell = 1}^{L-1}a_y^2[\ell]$; recall from Proposition~\ref{prop:gamma}:
\begin{equation} \label{eq:sigma2}
\sigma^2 = E_3 - (a^1_y)^2/\beta. 
\end{equation} 
Plugging into~\eqref{eq:E12} and rearranging, we get a first quadratic equation in $\beta$,
\begin{equation} \label{eq:quad1}
\mathcal{A}\beta^2 + \mathcal{B}\beta + \mathcal{C} = 0,
\end{equation}
where 
\begin{align*}
\mathcal{A} &= E_2 - (2L+1)a_y^1E_3, \\ 
\mathcal{B} &= -E_1 + (2L+1)(a_y^1)^3 + a_y^1E_3  , \\
\mathcal{C} &= -(a_y^1)^3.
\end{align*}
Importantly, these coefficients are observable quantities. As we assume throughout this proof that $x$ has nonzero mean, $a_y^1 \neq 0$ and we conclude that this equation is non-trivial.

Next, we derive the second quadratic equation for $\beta$. We notice that 
\begin{equation} \label{eq:E3}
E_4 = \frac{1}{L}(a_y^1)^3 = \frac{1}{L}\beta^3 (\one ^Tx)^3   = \beta^3 F_2,
\end{equation}
where $F_2 = \frac{1}{L}(\one ^Tx)^3$, and we can work out that:
\begin{equation*}
F_2 = a_x^3[0,0] + 3\sum_{j=1}^{L-1} \left(a_x^3[j,j] + a_x^3[0,j]\right) + 6\sum_{1\leq i < j\leq L-1}a_x^3[i,j].
\end{equation*}
Once again, $F_2$ can be estimated from $a_y^3$, taking bias and scaling into account:
\begin{align}
E_5 & = a_y^3[0,0] + 3\sum_{j=1}^{L-1} \left(a_y^3[j,j] + a_y^3[0,j]\right) + 6\sum_{1\leq i < j\leq L-1}a_y^3[i,j]  = L \beta F_2 + (6L-3)\sigma^2a_y^1.
\end{align}
Consider the following ratio:
\begin{equation*} 
\frac{E_5}{E_4} = \frac{L}{\beta^2} + \frac{(6L-3)\sigma^2a_y^1}{E_4}.
\end{equation*}
From the latter, we deduce:
\begin{equation*}
\sigma^2 = \frac{E_5}{a_y^1(6L-3)}  - \frac{LE_4}{\beta^2a_y^1(6L-3)}.
\end{equation*}
Using~\eqref{eq:sigma2} and rearranging, we get the second quadratic:
\begin{equation} \label{eq:quad2}
\mathcal{D}\beta^2 + \mathcal{E}\beta + \mathcal{F} = 0,
\end{equation}
where
\begin{align*}
\mathcal{D} &= E_3 - \frac{E_5}{a_y^1(6L-3)}, \\ 
\mathcal{E} &= -(a_y^1)^2, \\
\mathcal{F} &= \frac{LE_4}{a_y^1(6L-3)}.
\end{align*}
It is also non-trivial since $E_4 \neq 0$.
